Defining Transcriptome using RNA Sequencing\\ of {\em Streptococcus mutans}

### 2012-02-02 package/supplementary.Rnw
Just an update.

<<echo=FALSE,results=hide>>=
# not run when smutans R package is installed
# Use this when smutans R package is being tested.
library(DESeq)
library(goseq)
library(GenomicRanges)
library(plotrix)
library(xtable)
library(RColorBrewer)
source("smutans/R/class_and_slots.R")
source("smutans/R/core.R")
source("smutans/R/methods.R")
load("smutans/data/smutansData.tx.RData")
load("smutans/data/smutansData.txGenes.RData")
load("smutans/data/smutansData.txPileup.RData")
load("smutans/data/smutans.feature.genes.RData")
load("smutans/data/smutans.genes.criteria.RData")
load("smutans/data/smutans.go.genes.RData")
load("smutans/data/smutans.cat.desc.RData")
load("smutans/data/smutans.keg.genes.RData")
load("smutans/data/smutans.keg.desc.RData")
load("smutans/data/smutansGenes.RData")
@
<<eval=FALSE,echo=FALSE,results=hide>>=
# not run without smutans R package installed
# Remove eval=FALSE when the R package is installed
library(smutans)
load(smutansData.tx)
load(smutansData.txGenes)
load(smutansData.txPileup)
load(smutans.feature.genes)
data(smutans.genes.criteria)
data(smutans.go.genes)
data(smutans.cat.desc)
data(smutans.keg.genes)
data(smutans.keg.desc)
data(smutansGenes)
@

\section{Supplementary Text}
A
\subsection{Summary statistics for recombinant trees}
\citep{Gorke2008}

@@

%\section{Supplementary Tables}

<<echo=FALSE,results=hide>>=
geneDescription <- read.delim("feature-genome.out-geneonly.txt",head=FALSE,sep="\t")
geneDescription <- geneDescription[,c(4,7)]
colnames(geneDescription) <- c("Gene ID","Gene Description")

@@

<<echo=FALSE,results=tex>>=
#glucoseOnly <- newSmutans( smutansGenes, title="Glucose Only" )
#glucoseOnly <- smutans.de2( glucoseOnly, condition="glucose", 
#                            condA="ua159", condB="tw1" )
#save(glucoseOnly, file=file.path("glucoseOnly.RData"))
load("glucoseOnly.RData")
x <- smutans.de2List( glucoseOnly, qval=1e-3 )
x <- x[,c("id","baseMeanA","baseMeanB","log2FoldChange","padj")]
colnames(x) <- c("Gene ID","UA159","TW1","fold", "p-value")
x <- merge(x, geneDescription, by.x = "Gene ID", by.y = "Gene ID")
x.big <- xtable( x[order(x[,"p-value"]),],
                 display=c("s","s","d","d","f","e","s"),
                 digits=c(0,0,0,0,1,0,0),
                 align=c('l','l','r','r','r','r','p{3.5in}'),
                 label='tabGlucoseOnly',
                 caption='List of genes for the comparison UA159 and TW1 under glucose-grown cells',
               )
print( x.big,
       tabular.environment='longtable',
       caption.placement="top",
       include.rownames=FALSE,
       floating=FALSE) 
x <- merge(x, functionalDescription, by.x = "Gene ID", by.y = "Gene ID")
@
\clearpage{}

<<figPieGlucoseOnly,fig=TRUE,echo=FALSE,width=6,height=8>>=
par(mai=c(0,0,0,0),fig=c(0,1,0,1),mfrow=c(2,1))
#par(mfrow=c(2,1))
trim <- function (x) gsub("^\\s+|\\s+$", "", x)
y <- x[,"Functional Class"]
z <- lapply(y, function(x) trim(unlist(strsplit(as.character(x),"\\|\\|"))))
z <- table(as.factor(sort(unlist(z))))

@@

\clearpage{}
<<echo=FALSE,results=tex>>=
#galactoseOnly <- newSmutans( smutansGenes, title="Galactose Only" )
#galactoseOnly <- smutans.de2( galactoseOnly, condition="galactose", 
#                              condA="ua159", condB="tw1" )
#save(galactoseOnly, file=file.path("galactoseOnly.RData"))
load("galactoseOnly.RData")
x <- smutans.de2List( galactoseOnly, qval=1e-3 )
x <- x[,c("id","baseMeanA","baseMeanB","log2FoldChange","padj")]
colnames(x) <- c("Gene ID","UA159","TW1","fold", "p-value")
x <- merge(x, geneDescription, by.x = "Gene ID", by.y = "Gene ID")
x.big <- xtable( x[order(x[,"p-value"]),],
                 display=c("s","s","d","d","f","e","s"),
                 digits=c(0,0,0,0,1,0,0),
                 align=c('l','l','r','r','r','r','p{3.5in}'),
                 label='tabGalactoseOnly',
                 caption='List of genes for the comparison UA159 and TW1 under galactose-grown cells',
               )
print( x.big,
       tabular.environment='longtable',
       caption.placement="top",
       include.rownames=FALSE,
       floating=FALSE) 
x <- merge(x, functionalDescription, by.x = "Gene ID", by.y = "Gene ID")
@
\clearpage{}

<<figPieGalactoseOnly,fig=TRUE,echo=FALSE,width=6,height=8>>=
par(mai=c(0,0,0,0),fig=c(0,1,0,1),mfrow=c(2,1))
#par(mfrow=c(2,1))
trim <- function (x) gsub("^\\s+|\\s+$", "", x)
y <- x[,"Functional Class"]
z <- lapply(y, function(x) trim(unlist(strsplit(as.character(x),"\\|\\|"))))
z <- table(as.factor(sort(unlist(z))))

>>>>

<<echo=FALSE,results=hide>>=
# not run when smutans R package is installed
# Use this when smutans R package is being tested.
library(DESeq)
library(goseq)
library(GenomicRanges)
library(IRanges)
library(plotrix)
library(xtable)
library(RColorBrewer)
source("smutans/R/class_and_slots.R")
source("smutans/R/core.R")
source("smutans/R/methods.R")
load("smutans/data/smutansData.tx.RData")
load("smutans/data/smutansData.txGenes.RData")
load("smutans/data/smutansData.txGrCombined.RData")
load("smutans/data/smutansData.txAnnotation.RData")
load("smutans/data/smutansData.txRNAz.RData")
load("smutans/data/smutansData.txPileup.RData")
load("smutans/data/smutans.feature.genes.RData")
load("smutans/data/smutans.genes.criteria.RData")
load("smutans/data/smutans.go.genes.RData")
load("smutans/data/smutans.cat.desc.RData")
load("smutans/data/smutans.keg.genes.RData")
load("smutans/data/smutans.keg.desc.RData")
load("smutans/data/smutansGenes.RData")
@
<<eval=FALSE,echo=FALSE,results=hide>>=
# not run without smutans R package installed
# Remove eval=FALSE when the R package is installed
library(smutans)
load(smutansData.tx)
load(smutansData.txGenes)
load(smutansData.txCombined)
load(smutansData.txAnnotation)
load(smutansData.txRNAz)
load(smutansData.txPileup)
load(smutans.feature.genes)
data(smutans.genes.criteria)
data(smutans.go.genes)
data(smutans.cat.desc)
data(smutans.keg.genes)
data(smutans.keg.desc)
data(smutansGenes)
@

\section{Supplementary Text}
Add texts if needed.
\subsection{Subsection name}
Add texts if needed and citations such as \citep{Gorke2008}.

@@

\clearpage{}
\section{Supplementary Tables}

%\clearpage{}
<<echo=FALSE,results=tex>>=
ir1 <- gaps(ranges(smutansData.txAnnotation))
ir2 <- ranges(smutansData.txRNAz)
mtchUnannotatedRNA <- findOverlaps(ir2,ir1,type="within")
txRNAz.unannotated <- ir2[matchMatrix(mtchUnannotatedRNA)[,1]]
t1 <- data.frame(names=names(txRNAz.unannotated),start=start(txRNAz.unannotated),end=end(txRNAz.unannotated))
x.big <- xtable( t1,
                 display=c("s","s","d","d"),
                 align=c('l','l','r','r'),
                 label='tabUnannotatedRNA',
                 caption='{\\bf List of unannotated RNAs predicted using RNAz.}'
               )
print( x.big,
       caption.placement="top",
       include.rownames=FALSE)
@

\clearpage{}
%-----------------------------------------------------------------------------
\begin{table}[!h]
\caption{{\bf Gene Ontology enrichments for target genes of putative small RNAs.}}
\vspace{1ex}
\noindent \begin{centering}
\begin{tabular}{lcccl}
\hline 
small RNA & $q^a$ & count & GO & Description \\
\hline 
%-----------------------------------------------------------------------------
<<echo=FALSE,results=tex>>=
smutans.mannwhitney( qval=0.05,
                     genes.criteria=smutans.genes.criteria,
                     go.genes=smutans.go.genes,
                     cat.desc=smutans.cat.desc )
@
%-----------------------------------------------------------------------------
\hline 
\end{tabular}
\par\end{centering}
\begin{flushleft}
$^a$False discovery rate estimated by the
Benjamini-Hochberg method.  All categories having at least ten genes and
$q\leq 0.05$ are displayed.\\ 
\end{flushleft}
\label{tab:goseq}
\end{table}
%-----------------------------------------------------------------------------

<<echo=FALSE,results=hide>>=
trim <- function (x) gsub("^\\s+|\\s+$", "", x)

geneDescription <- read.delim("feature-genome.out-geneonly.txt",head=FALSE,sep="\t")
geneDescription <- geneDescription[,c(4,7)]
colnames(geneDescription) <- c("Gene ID","Gene Description")

@@

\clearpage{}
<<echo=FALSE,results=tex>>=
#glucoseOnly <- newSmutans( smutansGenes, title="Glucose Only" )
#glucoseOnly <- smutans.de2( glucoseOnly, condition="glucose", 
#                            condA="ua159", condB="tw1" )
#save(glucoseOnly, file=file.path("glucoseOnly.RData"))
load("glucoseOnly.RData")
x <- smutans.de2List( glucoseOnly, qval=1e-3 )
x <- x[,c("id","baseMeanA","baseMeanB","log2FoldChange","padj")]
colnames(x) <- c("Gene ID","UA159","TW1","fold", "p-value")
x <- merge(x, geneDescription, by.x = "Gene ID", by.y = "Gene ID")
x.big <- xtable( x[order(x[,"p-value"]),],
                 display=c("s","s","d","d","f","e","s"),
                 digits=c(0,0,0,0,1,0,0),
                 align=c('l','l','r','r','r','r','p{3.5in}'),
                 label='tabGlucoseOnly',
                 caption='{\\bf List of genes for the comparison UA159 and TW1 under glucose-grown cells.}',
               )
print( x.big,
       tabular.environment='longtable',
       caption.placement="top",
       include.rownames=FALSE,
       floating=FALSE) 
x <- merge(x, functionalDescription, by.x = "Gene ID", by.y = "Gene ID")
tableGlucoseOnly <- x
@

\clearpage{}
%-----------------------------------------------------------------------------
\begin{table}[!h]
\caption{{\bf Gene Ontology (GO) enrichments for differentially expressed
genes in the comparison UA159 and TW1 grown in glucose.}}
\vspace{1ex}
\noindent \begin{centering}
\begin{tabular}{cl}
\hline 
$q^a$ & Description \\
\hline 
%-----------------------------------------------------------------------------
<<echo=FALSE,results=tex>>=
smutans.de2Goseq ( glucoseOnly, qval=1e-3, 
                   feature.gene=smutans.feature.genes,
                   go.genes=smutans.go.genes,
                   cat.desc=smutans.cat.desc )
@
%-----------------------------------------------------------------------------
\hline 
\end{tabular}
\par\end{centering}
\begin{flushleft}
$^a$ We used a multiple-testing adjusted $p$-value of $1\times 10^{-3}$ to
determine differentially expressed genes.  For each category having at least ten
genes a variation of Fisher's exact test was performed, and another
multiple-testing adjusted $p$-value, or $q$-value was obtained. The listed
categories were chosen at the cutoff value of 0.05.
\\ 
\end{flushleft}
\label{tab:goseqGlucoseOnly}
\end{table}
%-----------------------------------------------------------------------------

\clearpage{}
<<echo=FALSE,results=tex>>=
#galactoseOnly <- newSmutans( smutansGenes, title="Galactose Only" )
#galactoseOnly <- smutans.de2( galactoseOnly, condition="galactose", 
#                              condA="ua159", condB="tw1" )
#save(galactoseOnly, file=file.path("galactoseOnly.RData"))
load("galactoseOnly.RData")
x <- smutans.de2List( galactoseOnly, qval=1e-3 )
x <- x[,c("id","baseMeanA","baseMeanB","log2FoldChange","padj")]
colnames(x) <- c("Gene ID","UA159","TW1","fold", "p-value")
x <- merge(x, geneDescription, by.x = "Gene ID", by.y = "Gene ID")
x.big <- xtable( x[order(x[,"p-value"]),],
                 display=c("s","s","d","d","f","e","s"),
                 digits=c(0,0,0,0,1,0,0),
                 align=c('l','l','r','r','r','r','p{3.5in}'),
                 label='tabGalactoseOnly',
                 caption='{\\bf List of genes for the comparison UA159 and TW1 under galactose-grown cells}',
               )
print( x.big,
       tabular.environment='longtable',
       caption.placement="top",
       include.rownames=FALSE,
       floating=FALSE) 
x <- merge(x, functionalDescription, by.x = "Gene ID", by.y = "Gene ID")
tableGalactoseOnly <- x
@

%\clearpage{}
%%-----------------------------------------------------------------------------
%\begin{table}[!h]
%\caption{{\bf Gene Ontology (GO) enrichments for differentially expressed
%genes in the comparison UA159 and TW1 grown in galactose.}}
%\vspace{1ex}
%\noindent \begin{centering}
%\begin{tabular}{cl}
%\hline 
%$q^a$ & Description \\
%\hline 
%%-----------------------------------------------------------------------------
%<<echo=FALSE,results=tex>>=
%smutans.de2Goseq ( galactoseOnly, qval=1e-3, 
%                   feature.gene=smutans.feature.genes,
%                   go.genes=smutans.go.genes,
%                   cat.desc=smutans.cat.desc )
%@
%%-----------------------------------------------------------------------------
%\hline 
%\end{tabular}
%\par\end{centering}
%\begin{flushleft}
%$^a$False discovery rate estimated by the
%Benjamini-Hochberg method.  All categories having at least ten genes and
%$q\leq 0.05$ are displayed.\\ 
%\end{flushleft}
%\label{tab:goseqGalactoseOnly}
%\end{table}
%%-----------------------------------------------------------------------------

\clearpage{}
<<echo=FALSE,results=tex>>=
#ua159only <- newSmutans( smutansGenes, title="UA159 Only" )
#ua159only <- smutans.de2( ua159only, type="ua159", 
#                          condA="glucose", condB="galactose" )
#save(ua159only, file=file.path("ua159only.RData"))
load("ua159only.RData")
x <- smutans.de2List( ua159only, qval=1e-10 )
x <- x[,c("id","baseMeanA","baseMeanB","log2FoldChange","padj")]
colnames(x) <- c("Gene ID","Glucose","Galactose","fold", "p-value")
x <- merge(x, geneDescription, by.x = "Gene ID", by.y = "Gene ID")
x.big <- xtable( x[order(x[,"p-value"]),],
                 display=c("s","s","d","d","f","e","s"),
                 digits=c(0,0,0,0,1,0,0),
                 align=c('l','l','r','r','r','r','p{3.5in}'),
                 label='tabUA159Only',
                 caption='{\\bf List of genes for the comparison of UA159 wild-type grown with glucose and galactose.} Note that we used 1e-10 as the p-value cutoff.',
               )
print( x.big,
       tabular.environment='longtable',
       caption.placement="top",
       include.rownames=FALSE,
       floating=FALSE) 
x <- merge(x, functionalDescription, by.x = "Gene ID", by.y = "Gene ID")
tableUA159Only <- x
@

\clearpage{}
<<echo=FALSE,results=tex>>=
#tw1only <- newSmutans( smutansGenes, title="TW1 Only" )
#tw1only <- smutans.de2( tw1only, type="tw1", 
#                          condA="glucose", condB="galactose" )
#save(tw1only, file=file.path("tw1only.RData"))
load("tw1only.RData")
x <- smutans.de2List( tw1only, qval=1e-3 )
x <- x[,c("id","baseMeanA","baseMeanB","log2FoldChange","padj")]
colnames(x) <- c("Gene ID","Glucose","Galactose","fold", "p-value")
x <- merge(x, geneDescription, by.x = "Gene ID", by.y = "Gene ID")
x.big <- xtable( x[order(x[,"p-value"]),],
                 display=c("s","s","d","d","f","e","s"),
                 digits=c(0,0,0,0,1,0,0),
                 align=c('l','l','r','r','r','r','p{3.5in}'),
                 label='tabTW1Only',
                 caption='{\\bf List of genes for the comparison of TW1 mutant-type grown with glucose and galactose.}',
               )
print( x.big,
       tabular.environment='longtable',
       caption.placement="top",
       include.rownames=FALSE,
       floating=FALSE) 
x <- merge(x, functionalDescription, by.x = "Gene ID", by.y = "Gene ID")
tableTW1Only <- x
@

\clearpage{}
%-----------------------------------------------------------------------------
\begin{table}[!h]
\caption{{\bf Gene Ontology (GO) enrichments for differentially expressed
genes in the comparison of glucose-grown and galactose-grown TW1 mutant-type
cells.}}
\vspace{1ex}
\noindent \begin{centering}
\begin{tabular}{cl}
\hline 
$q^a$ & Description \\
\hline 
%-----------------------------------------------------------------------------
<<echo=FALSE,results=tex>>=
smutans.de2Goseq ( tw1only, qval=1e-3, 
                   feature.gene=smutans.feature.genes,
                   go.genes=smutans.go.genes,
                   cat.desc=smutans.cat.desc )
@
%-----------------------------------------------------------------------------
\hline 
\end{tabular}
\par\end{centering}
\begin{flushleft}
$^a$False discovery rate estimated by the
Benjamini-Hochberg method.  All categories having at least ten genes and
$q\leq 0.05$ are displayed.\\ 
\end{flushleft}
\label{tab:goseqTW1Only}
\end{table}
%-----------------------------------------------------------------------------
\clearpage{}
%-----------------------------------------------------------------------------
\begin{table}[!h]
\caption{{\bf KEGG enrichments for differentially expressed
genes in the comparison of glucose-grown and galactose-grown TW1 mutant-type
cells.}}
\vspace{1ex}
\noindent \begin{centering}
\begin{tabular}{cl}
\hline 
$q^a$ & Description \\
\hline 
%-----------------------------------------------------------------------------
<<echo=FALSE,results=tex>>=
smutans.de2Goseq ( tw1only, qval=1e-3, 
                   feature.gene=smutans.feature.genes,
                   go.genes=smutans.keg.genes[c(4,3)],
                   cat.desc=smutans.keg.desc)
@
%-----------------------------------------------------------------------------
\hline 
\end{tabular}
\par\end{centering}
\begin{flushleft}
$^a$False discovery rate estimated by the
Benjamini-Hochberg method.  All categories having at least ten genes and
$q\leq 0.05$ are displayed.\\ 
\end{flushleft}
\label{tab:kegseqTW1Only}
\end{table}
%-----------------------------------------------------------------------------

% END OF TABLE
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\clearpage{}
\section{Supplementary Figures}

<<figClust,echo=FALSE,result=hide,fig=TRUE>>=
allSample <- newSmutans( smutansGenes )
smutans.de2Clust( allSample )
@
\begin{figure}
\centering
\includegraphics[width=.9\textwidth]{supplementary-figClust}
\caption{{\bf Cluster of 13 RNA-seq samples.} 
Heatmap shows the Euclidean distances between the samples
as calculated from the variance-stabilizing transformation of the
count data.}
\label{figClust}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Transcript analyses R scripts
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

<<figGrCombinedExpression,fig=TRUE,echo=FALSE>>=
l <- list(elementMetadata(smutansData.txGrCombined)$pileup)
bin <- c(as.integer(seq.int(0, 1000, length.out=15)), 50000)
multhist(l,breaks=bin)
@
\begin{figure}
\centering
\includegraphics[width=.9\textwidth]{supplementary-figGrCombinedExpression}
\caption{{\bf Average pileup values for segments.}
The last bin sums from 1000 bp to 5000 bp.
}
\label{figGrCombinedExpression}
\end{figure}

<<echo=FALSE,result=hide>>=
# library(GenomicRanges)
# library(IRanges)

notExpressedTx <- smutansData.txGrCombined[elementMetadata(smutansData.txGrCombined)$zero > 0.1 & elementMetadata(smutansData.txGrCombined)$pileup < 5]
grTx <- smutansData.txGrCombined[!(elementMetadata(smutansData.txGrCombined)$zero > 0.1 & elementMetadata(smutansData.txGrCombined)$pileup < 5)]

matchGenesTx <- findOverlaps(smutansData.txGenes,grTx)
matchAnnotationTx <- findOverlaps(smutansData.txAnnotation,grTx)

grNoAnnotation <- gaps(ranges(smutansData.txAnnotation))
grNoncoding <- gaps(ranges(smutansData.txGenes))
grRNAz <- ranges(smutansData.txRNAz)

matchUnannotatedRNA <- findOverlaps(grRNAz,grNoAnnotation,type="within")
txRNAz.unannotated <- grRNAz[matchMatrix(matchUnannotatedRNA)[,1]]

# A bug in GenomicRanges: the last segment is the while genome. I removed it
# from the noncoding ranges.
# matchNoncoding <- findOverlaps(grNoncoding,grTx)

# Expressed transcripts without annotated genes
grTxWGenes <- grTx[unique(matchMatrix(matchGenesTx)[,2])]
grTxWOGenes <- grTx[-unique(matchMatrix(matchGenesTx)[,2])]
grTxWOAnnotation <- grTx[-unique(matchMatrix(matchAnnotationTx)[,2])]

# Find transcripts with genes with conflicted strands
grGenesPlus <- smutansData.txGenes[strand(smutansData.txGenes)=='+']
grGenesMnus <- smutansData.txGenes[strand(smutansData.txGenes)=='-']
#mtchPlus <- findOverlaps(grGenesPlus,grTxWGenes)
#mtchMnus <- findOverlaps(grGenesMnus,grTxWGenes)
#txWithConflictedGenes <- intersect(matchMatrix(mtchPlus)[,2],matchMatrix(mtchMnus)[,2])

# Tx with genes strands
grTx2 <- grTxWGenes[!strand(grTxWGenes)=='*']

# Find the number of genes in segments.
matchGenesTx2 <- findOverlaps(smutansData.txGenes,grTx2,type="within")
number.genes.tx <- runLength(Rle(matchMatrix(matchGenesTx2)[,"subject"]))

# Length distribution of UTR regions
# grTx2 <- grTxWGenes[-txWithConflictedGenes] 
# Lengths of 5' and 3' UTRs
mtchPlus <- findOverlaps(grGenesPlus,grTx2,type="within")
mtchMnus <- findOverlaps(grGenesMnus,grTx2,type="within")
length5UTR <- c()
length3UTR <- c()
for (i in unique(matchMatrix(mtchPlus)[,"subject"])) {
  x <- matchMatrix(mtchPlus)[matchMatrix(mtchPlus)[,"subject"]==i,"query"]
  y <- min(start(grGenesPlus[x])) - start(grTx2[i])
  z <- end(grTx2[i]) - max(end(grGenesPlus[x])) 
  length5UTR <- c(length5UTR,y)
  length3UTR <- c(length3UTR,z)
}
for (i in unique(matchMatrix(mtchMnus)[,"subject"])) {
  x <- matchMatrix(mtchMnus)[matchMatrix(mtchMnus)[,"subject"]==i,"query"]
  y <- min(start(grGenesMnus[x])) - start(grTx2[i])
  z <- end(grTx2[i]) - max(end(grGenesMnus[x])) 
  length5UTR <- c(length5UTR,z)
  length3UTR <- c(length3UTR,y)
}

# Correlation of expression levels of adjacent genes.
pileup.x <- lapply(ranges(smutansData.txGenes), function(x) seqselect(smutansData.txPileup, x))
pileup.y <- mapply(function(x,y) sum(x)/y, x=pileup.x, y=width(smutansData.txGenes))
elementMetadata(smutansData.txGenes)["score"] <- pileup.y
mtch3 <- findOverlaps(smutansData.txGenes,grTx2)
x <- rle(matchMatrix(mtch3)[,"subject"])
comp.exp.two.genes <- c()
for (i in x$values[x$lengths > 1]) {
  y <- combn(matchMatrix(mtch3)[,"query"][matchMatrix(mtch3)[,"subject"]==i],2)
  for (j in 1:dim(y)[2]) {
    comp.exp.two.genes <- cbind(comp.exp.two.genes,as.matrix(elementMetadata(smutansData.txGenes)["score"][y[,j],]))
  }
}
# Correlation of expression levels of adjacent genes between transcripts.
comp.exp.two.genes.between.transcript <- c()
for (i in 1:length(x$values)) {
  if (i == length(x$values)) {
    break
  }
  y1 <- matchMatrix(mtch3)[,"query"][matchMatrix(mtch3)[,"subject"]==x$values[i]]
  i2 <- i + 1
  y2 <- matchMatrix(mtch3)[,"query"][matchMatrix(mtch3)[,"subject"]==x$values[i2]]
  l <- as.matrix(do.call(expand.grid, list(y1,y2)))
  for (j in 1:dim(l)[1]) {
    comp.exp.two.genes.between.transcript <- cbind(comp.exp.two.genes.between.transcript,as.matrix(elementMetadata(smutansData.txGenes)["score"][l[j,],]))
  }
}
@

<<figLength,fig=TRUE,echo=FALSE>>=
l <- list(expressed=width(smutansData.txGenes[matchMatrix(matchGenesTx)[,1]]), 
          notexpressed=width(smutansData.txGenes[-matchMatrix(matchGenesTx)[,1]]))
bin <- c(as.integer(seq.int(0, 3000, length.out=15)), 8500)
multhist(l,breaks=bin)
@
\begin{figure}
\centering
\includegraphics[width=.9\textwidth]{supplementary-figLength}
\caption{{\bf Length distribution for expressed and not-expressed genes.}
Darker gray
bars represent frequency of expressed genes, and lighter gray ones that of
not-expressed genes. The last bin sums from 3000 bp to 8500 bp.
}
\label{figLength}
\end{figure}

<<figTranscriptLength,fig=TRUE,echo=FALSE>>=
l <- list(coding=width(grTxWGenes), 
          noncoding=width(grTxWOGenes))
bin <- c(as.integer(seq.int(0, 10000, length.out=15)), 30000)
multhist(l,breaks=bin)
@
\begin{figure}
\centering
\includegraphics[width=.9\textwidth]{supplementary-figTranscriptLength}
\caption{{\bf Length distribution for expressed segments with annotated genes
and unannotated genes.}
Darker gray bars represent frequency of expressed segments with annotated
genes, and lighter gray ones that of without annotated genes that can be
potential non-coding regions. The last bin sums from 10000 bp to 30000 bp.
}
\label{figTranscriptLength}
\end{figure}

<<figUTRLength,fig=TRUE,echo=FALSE>>=
l <- list(five=length5UTR, three=length3UTR)
bin <- c(as.integer(seq.int(0, 300, length.out=19)), 6500)
multhist(l,breaks=bin)
@
\begin{figure}
\centering
\includegraphics[width=.9\textwidth]{supplementary-figUTRLength}
\caption{{\bf Length distribution for 5' and 3' UTR.}
Darker gray bars represent frequency of 5' UTR lengths, and lighter gray 3' UTR
lengths. The last bin sums from 300 bp to 6500 bp.
\fixme{Check the too long UTRs because they do not make sense.}
}
\label{figUTRLength}
\end{figure}

<<figExpressionTwoGene,fig=TRUE,echo=FALSE>>=
plot( log10(comp.exp.two.genes[1,]), log10(comp.exp.two.genes[2,]),
      xlab="Expression of first Gene in log10",
      ylab="Expression of second Gene in log10",
      pch=19,cex=0.5)
abline(b=1,lty=2)
@
<<figExpressionTwoGeneBetweenTranscript,fig=TRUE,echo=FALSE>>=
plot( log10(comp.exp.two.genes.between.transcript[1,]),
      log10(comp.exp.two.genes.between.transcript[2,]),
      xlab="Expression of first Gene in log10",
      ylab="Expression of second Gene in log10",
      pch=19,cex=0.5)
abline(b=1,lty=2)
@
\begin{figure}
\centering
\includegraphics[width=.6\textwidth]{supplementary-figExpressionTwoGene}
\includegraphics[width=.6\textwidth]{supplementary-figExpressionTwoGeneBetweenTranscript}
\caption{{\bf Scatter plot of expression levels of pairs of adjacent genes
located in the same transcripts.}
(Top) Two genes in the same tranascripts are plotted using their expression levels in
log10 scale.
(Bottom) Two genes in different tranascripts are plotted using their expression levels in
log10 scale.
}
\label{figExpressionTwoGene}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=0.3\textwidth]{figures/supplementary-figRNALocus101}
\includegraphics[width=0.3\textwidth]{figures/supplementary-figRNALocus7}
\includegraphics[width=0.3\textwidth]{figures/supplementary-figRNALocus68}
\includegraphics[width=0.3\textwidth]{figures/supplementary-figRNALocus8}
\includegraphics[width=0.3\textwidth]{figures/supplementary-figRNALocus10}
\includegraphics[width=0.3\textwidth]{figures/supplementary-figRNALocus69}
\includegraphics[width=0.3\textwidth]{figures/supplementary-figRNALocus1}
\caption{{\bf MFE structures drawing encoding positional entropy for
differentially expressed regions.}
Predicted non-coding RNAs are shown from top-left to bottom-right: locus101,
locus7, locus68, locus8, locus10, locus69, and locus1.
}
\label{figRNALocus1}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=0.3\textwidth]{figures/supplementary-figRNALocus99}
\includegraphics[width=0.3\textwidth]{figures/supplementary-figRNALocus103}
\includegraphics[width=0.3\textwidth]{figures/supplementary-figRNALocus105}
\includegraphics[width=0.3\textwidth]{figures/supplementary-figRNALocus106}
\includegraphics[width=0.3\textwidth]{figures/supplementary-figRNALocus107}
\includegraphics[width=0.3\textwidth]{figures/supplementary-figRNALocus108}
\includegraphics[width=0.3\textwidth]{figures/supplementary-figRNALocus110}
\includegraphics[width=0.3\textwidth]{figures/supplementary-figRNALocus2}
\caption{{\bf MFE structures drawing encoding positional entropy with expressed
regions.}
Predicted non-coding RNAs are shown from top-left to bottom-right: 
locus99, locus103, locus105, locus106, locus107, locus108, locus110, and locus2.
}
\label{figRNALocus2}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=0.3\textwidth]{figures/supplementary-figRNALocus5}
\includegraphics[width=0.3\textwidth]{figures/supplementary-figRNALocus6}
\includegraphics[width=0.3\textwidth]{figures/supplementary-figRNALocus7}
\includegraphics[width=0.3\textwidth]{figures/supplementary-figRNALocus8}
\includegraphics[width=0.3\textwidth]{figures/supplementary-figRNALocus52}
\includegraphics[width=0.3\textwidth]{figures/supplementary-figRNALocus68}
\includegraphics[width=0.3\textwidth]{figures/supplementary-figRNALocus69}
\includegraphics[width=0.3\textwidth]{figures/supplementary-figRNALocus85}
\caption{{\bf MFE structures drawing encoding positional entropy with expressed
regions.}
Predicted non-coding RNAs are shown from top-left to bottom-right: 
locus5, locus6, locus7, locus8, locus52, locus68, locus69, and locus85.
}
\label{figRNALocus3}
\end{figure}

\clearpage{}
\begin{figure}
\centering
\includegraphics[width=\textwidth]{figures/ucsc}
\caption{{\bf UCSC genome browser tracks of RNA-seq data.}
Tracks are shown from the top: known genes, pileup value tracks for UA159 grown
in glucose, those for TW1 in glucose, and the five segments predicted based on
RNA-seq data and known genes. Known genes with forward strand are colored red,
and those of reverse strand blue. The most top of the five predicted segments
are also colored in the same way. 
}
\label{figUCSC}
\end{figure}

\clearpage{}
\begin{figure}
\centering
\includegraphics[width=\textwidth]{figures/ucsc2}
\caption{{\bf UCSC genome browser tracks of RNA-seq data for \textit{galKTE} genes.}
Four genes from SMU.886 to SMU.889 were also upregulated in respose to
galactose. RNA-seq1 is UA159 with glucose, RNA-seq2 TW1 with glucose,
RNA-seq21 UA159 with galactose, and RNA-seq3 TW1 with galactose.
}
\label{figUCSC2}
\end{figure}

\clearpage{}
<<figPieGlucoseOnly,fig=TRUE,echo=FALSE,width=6,height=8>>=
par(mai=c(0,0,0,0),fig=c(0,1,0,1),mfrow=c(2,1))
y <- tableGlucoseOnly[,"Functional Class"]
z <- lapply(y, function(x) trim(unlist(strsplit(as.character(x),"\\|\\|"))))
z <- table(as.factor(sort(unlist(z))))

@@

<<figDotplotGlucoseOnly,fig=TRUE,echo=FALSE,width=12,height=6>>=
#par(mar=c(4.1, 3.1, 0.1, 0.1))
par(mai=c(0, 0, 0, 0))
y <- tableGlucoseOnly[,"Functional Class"]
# z <- lapply(y, function(x) trim(unlist(strsplit(as.character(x),"\\|\\|"))))
# z <- table(as.factor(sort(unlist(z))))
z <- sapply(y, function(x) {x1=strsplit(as.character(x),"\\|\\|"); lapply(x1, trim)})
xm <- tableGlucoseOnly[rep(seq(length(z)),sapply(z,length)),]
xm[,"Functional Class"] <- unlist(z)
# Use xm to plot a dotplot.
dotplot(xm[,"Functional Class"]~xm[,"fold"],
        main="",xlab="Log2 fold change",col="black",pch=88)
@
\begin{sidewaysfigure}
\centering
\includegraphics[width=0.9\textwidth]{supplementary-figDotplotGlucoseOnly}
\caption{{\bf Distribution of functional classes of differntially expressed genes
in comparing UA159 wild-type and TW1 mutant-type grown in glucose.}
}
\label{figDotplotGlucoseOnly}
\end{sidewaysfigure}

<<figFoldPValueGlucoseOnly,fig=TRUE,echo=FALSE>>=
par(mar=c(4.1, 4.1, 0.2, 0.2))
plot(abs(xm[,"fold"])~log10(xm[,"p-value"]),
     xlab="log10 p-value",ylab="Absolute log2 fold change",
     pch=19,cex=0.7)
@
\begin{figure}
\centering
\includegraphics[width=0.9\textwidth]{supplementary-figFoldPValueGlucoseOnly}
\caption{{\bf Scatter plot of fold versus p-vaule in log scale.}
}
\label{figFoldPValueGlucoseOnly}
\end{figure}

\clearpage{}
<<figPieGalactoseOnly,fig=TRUE,echo=FALSE,width=6,height=8>>=
par(mai=c(0,0,0,0),fig=c(0,1,0,1),mfrow=c(2,1))
y <- tableGalactoseOnly[,"Functional Class"]
z <- lapply(y, function(x) trim(unlist(strsplit(as.character(x),"\\|\\|"))))
z <- table(as.factor(sort(unlist(z))))
